Tropical Andes Taxonomic Diversity of Plants and Frugivores

### 2023-12-13 code/L2/L2_1_TropicalAndes_taxdiv.Rmd
add 100 km resolution and code to plot r-squared values by resolution

# Compare r-squared by resolution
Plot
# Multi panel plots
Richness maps
# Write data to csv
```{r}
write.csv(plant_cellRichness0.75, file.path(output_path,"TropicalAndes_GBIF_plant_richness_sub.csv"))
write.csv(frugivore_cellRichness0.75, file.path(output_path,"TropicalAndes_frugivore_richness_sub.csv"))
```

>>>>

```{r}
plot(plant_cellRichness_75km$num_species, frugivore_cellRichness_75km$num_species,
     main = "Plant Richness vs Frugivore Richness [75 km]",
     xlab = "Plant Richness by cell", ylab = "Frugivore Richness by cell")
abline(lm(frugivore_cellRichness_75km$num_species ~ plant_cellRichness_75km$num_species))
```

```{r}
hist(plant_cellRichness_75km$num_species,
     main = "Histogram of Plant Richness by cell",
     xlab = "Plant Richness by cell")
hist(frugivore_cellRichness_75km$num_species,
     main = "Histogram of Frugivore Richness by cell",
     xlab = "Frugivore Richness by cell")
```

## 100 km 
### Make Grid
```{r}
TAGrid_100 <- TApoly %>%
  st_make_grid(cellsize = c(100000)) %>%
  st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
  st_cast("MULTIPOLYGON") %>%
  st_sf() %>%
  mutate(cellid = row_number())
```

### Plot Empty Grid
```{r}
gridPlot <-
ggplot() +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TropicalAndes_IUCNHabitat_Forest) +
  geom_sf(data = TAGrid_100) +
  labs(title = "Tropical Andes Grid [100 km]", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

### Cell Richness
```{r}
plant_richness_grid_100 <- TAGrid_100 %>%
  st_join(plants_sf_species) %>%
  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
  group_by(cellid) %>%
  summarize(num_species = sum(overlap))
```

```{r}
frugivore_richness_grid_100 <- TAGrid_100 %>%
  st_join(frugivores_sf_species) %>%
  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
  group_by(cellid) %>%
  summarize(num_species = sum(overlap))
```

### Richness plots
```{r}
plantgridRichTA_100km <-
ggplot(plant_richness_grid_100) +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_sf(aes(fill = num_species), color = NA) +
  scale_fill_scico(palette = "romaO", direction = 1) +
  labs(title = "Tropical Andes Plant Richness [100 km]", x = "Longitude", y = "Latitude",
       fill = "Number of species") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

plantgridRichTA_100km
```

```{r}
frugivoregridRichTA_100km <-
ggplot(frugivore_richness_grid_100) +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_sf(aes(fill = num_species), color = NA) +
  scale_fill_scico(palette = "romaO", direction = 1) +
  labs(title = "Tropical Andes Frugivore Richness [100 km]", x = "Longitude", y = "Latitude",
       fill = "Number of species") +
    coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

frugivoregridRichTA_100km
```

## Extract cell values
```{r}
plant_cellRichness_100km <- plantgridRichTA_100km[["data"]]
frugivore_cellRichness_100km <- frugivoregridRichTA_100km[["data"]]
```

### Plant vs frugivore richness plot
```{r}
compare_richness_summary_100km <- summary(lm(frugivore_cellRichness_100km$num_species ~ plant_cellRichness_100km$num_species))
compare_richness_summary_100km

# save r-squared value
compare_richness_r2_100km <- compare_richness_summary_100km$r.squared
```

```{r}
plot(plant_cellRichness_100km$num_species, frugivore_cellRichness_100km$num_species,
     main = "Plant Richness vs Frugivore Richness [100 km]",
     xlab = "Plant Richness by cell", ylab = "Frugivore Richness by cell")
abline(lm(frugivore_cellRichness_100km$num_species ~ plant_cellRichness_100km$num_species))
```

```{r}
hist(plant_cellRichness_100km$num_species,
     main = "Histogram of Plant Richness by cell",
     xlab = "Plant Richness by cell")
hist(frugivore_cellRichness_100km$num_species,
     main = "Histogram of Frugivore Richness by cell",
     xlab = "Frugivore Richness by cell")
```

# Compare r-squared by resolution
Plot
```{r}
# create dataframe of resolutions and r-squared values
resolution <- c("5", "25", "50", "75", "100")
r_squared <- c(compare_richness_r2_5km, compare_richness_r2_25km, compare_richness_r2_50km, compare_richness_r2_75km, compare_richness_r2_100km)
r_squared_compare <- data.frame(resolution, r-squared)
```

```{r}
ggplot(r_squared_compare, aes(x=resolution, y=r_squared)) +
  geom_point()+
  labs(title = "R-squared values of plant-frugivore richness relationships by resolution", x = "Resolution (km)", y = "R-squared",
       fill = "Number of species") 
```

# Multi panel plots
Richness maps
# Write data to csv
```{r}
write.csv(plant_cellRichness0.75, file.path(output_path,"TropicalAndes_GBIF_plant_richness_sub.csv"))
write.csv(frugivore_cellRichness0.75, file.path(output_path,"TropicalAndes_frugivore_richness_sub.csv"))
```
